The Sub-Millimetre User Reduction Facility: an iterative map-maker for \scuba

### 2012-06-13 smurf_paper.tex
more intro work

\begin{abstract}
  We describe the properties of data from the Submillimetre Common
  User Bolometer Array 2 (\scuba), a camera for the 15-m James Clerk
  Maxwell Telescope that provides simultaneous imaging capabilities at
  450 and 850\,\micron, and the production of maps using the
  Sub-Millimetre User Reduction Facility (SMURF). With nearly 5000
  bolometers in each imaging band, and an approximately 5\,arcmin
  field-of-view, \scuba\ is presently the largest submillimetre camera
  in the world, and produces data at rates an order-of-magnitude
  faster than previous instruments. We have adopted a fast, iterative
  approach to map-making that enables data reduction on single,
  modern, high-end desktop computers with execution times that are
  shorter than the observing times.  SMURF is used in an automated
  setting both at the telescope for real-time feedback to observers,
  as well as for the production of generic science products at the
  Canadian Astronomical Data Centre. A suite of tools are also
  provided for interactive data analysis.  Three detailed case studies
  are used to: (i) explore convergence properties of the map-maker
  using simple prior constraints (Uranus -- a point source); (ii)
  demonstrate that our strategy is capable of recovering angular
  scales up to the size of the array footprint (approximately
  5\,arcmin) for bright extended sources (star-forming region M17);
  and (iii) achieve the white-noise limit of the instrument for faint
  point-source studies (blank-field survey of the Lockman Hole).
\end{abstract}

@@

Submillimetre imaging cameras generally use bolometers to maximize
sensitivity, in which an absorber collects photons that pass through
broad-band filters, and the signal is read out-out using a thermometer
\citep[e.g.][]{griffin2002}. The sensitivity of bolometers in a camera
like \scuba\ are limited by photon noise from the instrument and
ambient backgrounds. The low-frequency noise, however, is dominated by
sources such as slow variations in mean backgrounds (e.g., thermal
variations within the cryostat, and the atmosphere for ground-based
cameras), and drift in the readout electronics. Such noise has a power
spectrum $\propto 1/f^\alpha$ ($\alpha>0$), and the frequency at which
it is comparable to the white noise level is called the ``$1/f$
knee''. Since the low-frequency noise is largely correlated between
all, or subsets, of the bolometers in time, it can be suppressed
during map-making, since astronomical signals have the distinct
property that they are fixed in a sky reference frame (assuming they
are not time-varying). If successful, the noise in the resulting map
is said to be ``background noise limited'', meaning that the map noise
is uncorrelated spatially, and has an amplitude that scales as the
$\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the noise-equivalent
flux-density (the white noise level of a bolometer in 1\,s of
integration), and $t$ is the amount of integration time in a map
pixel.

There are numerous ways to attack the problem of map production, both
in terms of the data-collection method, and processing. The two most
important principles to follow in terms of scan strategy are: (i) to
modulate the astronomical signals of interest in such a way that they
appear in the lowest-noise regions of the bolometer noise power
spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
``cross-linking'', in which each portion of the map is sampled on a
range of temporal scales, again, to help distinguish time-varying
noise features from fixed astronomical sources. In the case of \scuba,
(i) is achieved through fast-scanning of the entire telescope (up to
600\,arcsec\,sec$^{-1}$ slew speeds); and (ii) by offering scan
patterns that cross the sky at a wide range of position angles,
ultimately observing each portion of the map at a number of different
times. Such methods are now used by virtually all existing
ground-based submillimetre cameras, in favour of ``chopping'' methods
(where the secondary is quickly moved to modulate the signal) that
were more appropriate for older instruments that had poorer
low-frequency noise performance, and were only sensivitve to modest
angular scales
\citep[e.g.][]{glenn1998,weferling2002,laurent2005,wilson2008,kovacs2008b}.

There are three general styles of map-making that are described in the
literature for bolometer data. The best maps are produced using
\emph{maximum likelihood} techniques, in which the map is expressed as
a linear combination of the data using weights that are a function of
the time-domain noise properties of the bolometers. The first good
description of this method appears in \citet{janssen1992} for the
production of maps from the COsmic Background Explorer
(COBE). However, this method can be extremely computationally
expensive, even when small numbers of bolometers are involved,
although we note that an algorithm called ``SANEPIC'' for reducing
maps from the Balloon-borne Large-Aperture Submillimeter Telescope
(BLAST) involving hundreds of bolometers and correctly incorporating
inter-bolometer noise correlations, has been implemented with
reasonable execution speeds and memory requirements
\citep[][]{patanchon2008}.

  Styles of map-making: least
squares \citep{janssen1992} maximum likelihood
\citep[e.g.,][]{patanchon2008}; iterative -- see \citet{johnstone2000}
implementation of \citet{wright1996} to remove chop,
\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA analysis
\citep[e.g.][]{laurent2005,scott2008,aguirre2010}.

>>>>

\begin{abstract}
  We describe data properties of the Submillimetre Common User
  Bolometer Array 2 (\scuba), a camera for the 15-m James Clerk
  Maxwell Telescope that provides simultaneous imaging capabilities at
  450 and 850\,\micron, and the production of maps using the
  Sub-Millimetre User Reduction Facility (SMURF). With nearly 5000
  bolometers in each imaging band, and an approximately 5\,arcmin
  field-of-view, \scuba\ is presently the largest submillimetre camera
  in the world, and produces data at rates an order-of-magnitude
  faster than previous instruments. We have adopted a fast, iterative
  approach to map-making that enables data reduction on single,
  modern, high-end desktop computers with execution times that are
  shorter than the observing times.  SMURF is used in an automated
  setting both at the telescope for real-time feedback to observers,
  as well as for the production of generic science products at the
  Canadian Astronomical Data Centre. A suite of tools are also
  provided for interactive data analysis.  Three detailed case studies
  are used to: (i) explore convergence properties of the map-maker
  using simple prior constraints (Uranus -- a point source); (ii)
  demonstrate that our strategy is capable of recovering angular
  scales up to the size of the array footprint (approximately
  5\,arcmin) for bright extended sources (star-forming region M17);
  and (iii) achieve the white-noise limit of the instrument for faint
  point-source studies (blank-field survey of the Lockman Hole).
\end{abstract}

@@

Submillimetre imaging cameras generally use bolometers to maximize
sensitivity, in which an absorber collects photons that pass through
broad-band filters, and the signal is read out-out using a thermometer
\citep[e.g.][]{griffin2002}. The sensitivity of bolometers in a camera
like \scuba\ are limited by photon noise from the instrument and
ambient backgrounds. The low-frequency noise, however, is dominated by
sources such as slow variations in mean backgrounds (e.g., thermal
variations within the cryostat, and the atmosphere for ground-based
cameras), and drift in the readout electronics. Such noise has a power
spectrum $\propto 1/f^\alpha$ ($\alpha>0$), and the frequency at which
it is comparable to the white noise level is called the ``$1/f$
knee''. Since the low-frequency noise is largely correlated between
all, or subsets, of the bolometers in time, it can be suppressed
during map-making, since astronomical signals have the distinct
property that they are fixed in a sky reference frame (assuming they
are not time-varying). If successful, the noise in the resulting map
is said to be ``white noise limited'', meaning that the map noise is
uncorrelated spatially, and has an amplitude that scales as the
$\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the noise-equivalent
flux-density (the white noise level of a bolometer in 1\,s of
integration), and $t$ is the amount of integration time in a map
pixel.

There are numerous ways to attack the problem of map production, both
in terms of the data-collection method, and processing. The two most
important principles to follow in terms of scan strategy are: (i) to
modulate the astronomical signals of interest in such a way that they
appear in the lowest-noise regions of the bolometer noise power
spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
``cross-linking'', in which each portion of the map is sampled on a
range of temporal scales, again, to help distinguish time-varying
noise features from fixed astronomical sources. In the case of \scuba,
(i) is achieved through fast-scanning of the entire telescope (up to
600\,arcsec\,sec$^{-1}$ slew speeds), such that significant drift in
the bolometers due to low-frequency noise occurs more slowly than the
crossing times for the astronomical scales of interest; and (ii) by
offering scan patterns that cross the sky at a wide range of position
angles, ultimately observing each portion of the map at a number of
different times. Such methods are now used by virtually all existing
ground-based submillimetre cameras, in favour of ``chopping'' methods
(where the secondary is quickly moved to modulate the signal) that
were more appropriate for older instruments that had poorer
low-frequency noise performance, and were only sensivitve to modest
angular scales
\citep[e.g.][]{glenn1998,weferling2002,wilson2008,kovacs2008b}.

There are three general styles of map-making that are relevant to
reducing bolomter data in the literature. The simplest methods involve
some basic processing of the data to remove as much noise
contamination as possible, e.g., using baseline removal, filters, and
then simply re-gridding the data into a map. Such was the basic recipe
for the baseline reduction of chopped data from SCUBA-2's predecessor
SCUBA (reference?), and MAMBO, another camera from the same
generation. Generally speaking, such methods are fast, although
dependending on the science goals and noise properties of the data,
they may not achieve the best theoretical noise performance on the
angular scales of interest. A particular case of this method was
successfully used for the reduction of data from BOLOCAM
\citep[e.g.][]{laurent2005}, and its younger sibling AzTEC
\citep[e.g.][]{scott2008}, using ``principal component analysis''
(PCA) cleaning. A statistical black-box is used to remove the most
correlated components of the bolometer signals, enabling the detection
of point-sources very close to the theoretical white-noise limits of
the detectors with reasonable computation times. However, PCA cleaning
is not a good generic solution for producing maps of extended
structures, since large astronomical sources produce correlated
signals amongst many detectors, and are removed by this procedure.

The best map-making strategies for recovering information on all
angular scales are \emph{maximum likelihood} techniques, in which the
time-series data are expressed as a sampling of the ``true'' map of
the sky plus noise, and then the equation is inverted to estimate the
map as some weighted linear combination of the data that minimizes the
variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the CMB literature include
\citet{tegmark1997} and \citet{stompor2002}.  The downside to such
methods is that they can be both computationally expensive and have
large memory requirements. While for some experimental designs fast
iterative methods for the inversion do exist with minimal memory usage
\citep[e.g.][]{wright1996}, for the more general map-making problem
involving many detectors, the problem is significantly complicated
both by the need to measure the cross power-spectra for all unique
pairs of bolometers, and then to perform the inversion itself. The
most promising maximum-likelihood method that may one day be applied
to SCUBA-2 data is ``SANEPIC'', which successfully reduced maps from
the Balloon-borne Large-Aperture Submillimeter Telescope, involving
hundreds of bolometers, and correctly incorporating inter-bolometer
noise correlations \citep[][]{patanchon2008}.

%  Styles of map-making: least
%squares \citep{janssen1992} maximum likelihood
%\citep[e.g.,][]{patanchon2008}; iterative -- see \citet{johnstone2000}
%implementation of \citet{wright1996} to remove chop,
%\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA analysis
%\citep[e.g.][]{laurent2005,scott2008,aguirre2010}.
